---
title: '**First Part: COVID-19 in Spain**'
subtitle: '**Advanced Regression And Prediction**'
author: '*Roberto Jesús Alcaraz Molina*'
date: "09/05/2021"
output:
  pdf_document:
    number_section: true
    highlight: tango
    fig_caption: yes
    toc: true
    # toc_depth: 4
header-includes: 
- \usepackage{float}
- \usepackage{amsbsy}
- \usepackage{amsmath}
- \usepackage{graphicx}
- \usepackage{subfig}
- \usepackage{booktabs}
bibliography: references.bib
---

```{=tex}
\begin{center}
\includegraphics[width=3in]{logouc3m}
\end{center}
```
\newpage

```{r, echo = F}
pacman::p_load(COVID19, tidyverse, tidymodels)

# spain <- COVID19::covid19(country = "spain")
# saveRDS(spain, "spain.RDS")
```

```{r include=FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      echo = F, 
                      message = FALSE,
                      eval = F,
                      fig.pos="H", 
                      fig.align="center",
                      fig.width=15,
                      cache=FALSE, error = TRUE)
```

# Introduction

The aim of this project is to analyze the coronavirus pandemic (COVID-19) in Spain. The data has been taken from the R package `COVID19` [@guidotti2020].

```{r}
spain <- readRDS("spain.RDS")
```

# Data Cleaning 

https://covid19datahub.io/articles/doc/data.html

The Spain COVID-19 data set contains 451 observations, which are the days from the 22nd of January 2020 until now, and 36 variables divided in 5 categories: *identifiers*, *COVID-19 variables*, *policy measures*, *geographic information* and *external keys*.

Regarding the *identifiers*, since we will only analyze Spain, we will remove the variable `id`, which denote the country, and will select the `date`.

In the second group, we have variables related with the COVID-19 such as the cumulative number of `deaths`, `confirmed` cases, `test`, etc.

```{r}
spain %>%
  select(-id)
```



### Identifiers:
- `id`:
- `date`:

### COVID-19 variables:
- `deaths`:
- `confirmed`:
- `test`:
- `vaccines`:
- `recovered`:
- `hosp`:
- `icu`:
- `vent`:
- `population`:

### Policy measures:
- `school_closing`:
- `workplace_closing`:
- `cancel_events`:
- `transport_closing`:
- `stay_home_restrictions`:


```{r}

```



# Exploratoy Data Analysis (EDA)


# Feature Engineering 


# Model tuning and selection


# References
